Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-11-19 grove-paper.tex
finish the decomposition algorithm
also various minor edits

A graph $\G : \E \rightarrow \Sigma$ is a function from edges to edge states,
where $\E = \U \times \V \times \P \times \V$,
unique IDs are drawn from some suitable set $\U$,
vertices are drawn from $\V$,
positions are drawn from $\P$,
and edge states are drawn from $\Sigma$, all defined below.

Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$,
and connects the child position $p$ of a source vertex $v$ to a target vertex $v^\prime$.
The state of each edge, $\e$, is determined by the corresponding edge state, $\G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$.
If $\G(\e) = \bot$, then $\e$ does not yet exist.
If $\G(\e)$ is $\Plus$, the edge has been created.
If $\G(\e)$ is $\Minus$, the edge has been destroyed.
The total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.

A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k$,
where $k \in \K$ corresponds to one of the forms defined in \autoref{fig:Syntax
as a grammar} that is not an operator or a hole. Each contructor has an arity,
which we define as a set of positions. A position $p \in \P$ identifies a
location, adjacent to some parent vertex, from which edges to child vertices may
originate. The $\arity$ function in \autoref{fig:Constructors, Indexes and
Arity} gives the complete mapping from constructors to potential edge
positions. Every graph contains a distinguished root vertex, root constructor,
and root position, respectively named $v_0$, $k_0$, and $p_0$.

A graph can be derived from an expression in the calculus by starting with an
empty graph, selecting unique ids $u, u_k$ and constructor $k$ for the outermost
expression, and creating edge $\e = (u, v_0, p_0, (u_k, k))$. Then, for each
sub-term at some position $p \in \arity(k)$, select unique id $u'$ and
constructor $k'$ and create edge $e' = (u', v, p, v')$ with $v = (u_k, k)$, $v'
= (u', k')$. Continuing in this manner, treat each sub-term, starting with the
ones at positions in $\arity(k')$, as if it were the outermost and create edges
to the vertices representing its immediate sub-terms until only operators and
holes remain. Note that empty holes are not modeled explicitly in the graph.
Wherever an empty hole term appears, the graph has no edges originating from
that position.

@@

  \DontPrintSemicolon
  \SetAlgoNoLine
  \SetKwFunction{reachable}{reachable}
  \SetKwFunction{simpleCycles}{simpleCycles}
  \SetKwProg{Fn}{function}{}{end}

  \KwIn{Graph $G$}
  \KwOut{Decomposition of $G$ into cycle-free expressions}

  $liveEdges := \{\e : G(\e) = \Plus\}$\;
  $liveVertexes := \bigcup_{(u, v, p, v') \in liveEdges} \{v, v'\}$\;
  $allVertexes := \bigcup_{(u, v, p, v') \in \dom(G)} \{v, v'\}$\;
  $multiparented := \{v \in allVertexes : \exists (u_1, v_1, p_1, v), (u_2, v_2, p_2, v) \in \dom(G), v_1 \ne v_2\}$\;
  $deleted := allVertexes \setminus liveVertexes \setminus \{v_0\}$\;
  $MP := \{\reachable(v, \varnothing) : v \in multiparented\}$\;
  $D := \{\reachable(v, \varnothing) : v \in deleted\}$\;
  \Return{$(\reachable(v_0, \varnothing), MP, D, \simpleCycles(deleted))$}\;
  \;

  \Fn{\reachable $((u, k), seen)$}{
    \If{$(u, k) \in multiparented \cup seen$}{
      \Return{$u$}
    }
    $childEdges := \{(u', v, p, v') \in liveEdges : v = (u, k)\}$\;
    $children := \{(p, \reachable(v, seen \cup \{u\})) : (p, v) \in childEdges\}$\;
    \Return{$(k, children)$}
  }
  \;

  \Fn{\simpleCycles $(remaining)$}{
    \Return{}
  }
  \;
\end{algorithm}

>>>>

A graph $G : \E \rightarrow \Sigma$ is a function from edges to edge states,
where $\E = \U \times \V \times \P \times \V$,
unique IDs are drawn from some suitable set $\U$,
vertices are drawn from $\V$,
positions are drawn from $\P$,
and edge states are drawn from $\Sigma$, all defined below.

Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$,
and connects the child position $p$ of source vertex $v$ to target vertex $v^\prime$.
The state of each edge, $\e$, is determined by the corresponding edge state, $G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$.
When $G(\e) = \bot$, $\e$ has not been created yet.
When $G(\e)$ is $\Plus$, the edge has been created.
When $G(\e)$ is $\Minus$, the edge has been destroyed.
The total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.

A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k \in
\K$, corresponding to a form defined in \autoref{fig:Syntax as a grammar} that
is neither an operator nor a hole. Each contructor has an arity, which we define
as a set of positions. A position $p \in \P$ identifies a location, adjacent to
some source vertex, from which edges to target vertices may originate. The
$\arity$ function in \autoref{fig:Constructors, Indexes and Arity} gives the
complete mapping from constructors to potential edge positions. Every graph
contains a distinguished root vertex, root constructor, and root position,
respectively named $v_0$, $k_0$, and $p_0$.

% TODO: should graph "recomposition" have an algorithm, too?
A graph can be derived from an expression in the calculus by starting with an
empty graph, selecting unique ids $u, u_k$ and constructor $k$ for the outermost
expression, and creating edge $\e = (u, v_0, p_0, (u_k, k))$. Then, for each
sub-term at some position $p \in \arity(k)$, select unique id $u'$ and
constructor $k'$ and create edge $e' = (u', v, p, v')$ with $v = (u_k, k)$, $v'
= (u', k')$. Continuing in this manner, treat each sub-term, starting with the
ones at positions in $\arity(k')$, as if it were the outermost and create edges
to the vertices representing its immediate sub-terms until only empty holes
remain. Note that empty holes are not modeled explicitly in the graph. Wherever
an empty hole appears in an expression, the corresponding graph position has no
edges originating from it.

@@

  \DontPrintSemicolon
  \SetAlgoNoLine
  \SetKwFunction{reachable}{reachable}
  \SetKwProg{Fn}{function}{}{end}

  \KwIn{Graph $G$}
  \KwOut{A partitioning of the edges of $G$}

  $live := \{\e : G(\e) = \Plus\}$\;
  $R := \{\reachable(\e) : \e = (u, v_0, p, v) \in live\}$\;
  $D := \{\reachable(\e) : \e \notin live\}$\;
  $muliparent := \{(u,v,p,v') \in live : \exists (u_1, v_1, p_1, v), (u_2, v_2, p_2, v) \in live, v_1 \ne v_2\}$\;
  $MP := \{\reachable(\e) : \e \in multiparent\}$\;
  $SC := \varnothing$\;
  $remaining := \{\e \notin \edges(R \cup D \cup MP)\}$\;
  \While{$remaining \ne \varnothing$}{
    $(root, cycle) := \reachable(\min(remaining))$\;
    $remaining := remaining \setminus \{root\} \setminus cycle$\;
    $SC := SC \cup \{(root, cycle)\}$\;
  }
  \Return{$(R, D, MP, SC)$}\;
  \;

  \Fn{$\reachable\ ((u, v, p, v'))$}{
    $children := \{(u', v', p', v'') \in live\}$\;
    $descendants := \bigcup_{\e \in children} \reachable(\e)$\;
    \Return{$((u, v, p, v'), descendants)$}
  }
\end{algorithm}
